On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-06-05 reports/white_paper/formal_glossary.tex
minor changes

\Paragraph{Attacker as a Dis-honest Agent}
As we mentioned in Section~\ref{sec:problem} when we postulated the
security hypothesis, Weaknesses and Vulnerability are the two core security
concepts of the current understanding of what security is. However, we see two
(opposed in terms of abstraction) possibility to better defined those concepts:
defining how an agent (Mankind or Nature) can maliciously use Weaknesses to
produce Vulnerability (and, in turn, Exploits), or how Errors can be defined
(since Error is a term from which Weakness are, in turn, defined).  The former,
in our current formalization, implies detailing and constraining the Causality
Principle by introducing constraints on how agents acts. For example,
introducing the concept of intentionality, motivation, or desire (see the BDI
framework defined in \autocite{Rao1995semantics} for a general overview of those concepts).
The symbolic approach (also called Dolev-Yao model) in the formal verification
of security and communication protocols, can be used in this spirit. The
attacker can be seen as an additional agent (often identified with the network
where the messages are exchanged) that sees the network traffic and manipulates
it. The actions are generic (the attacker can combine two messages, split a
message into its components, or decrypt/encrypt if it has the proper security
key) but not something that an honest agent cannot do. The main difference is
that an honest agent follows the nominal specification of the protocol while
the attacker doesn't. The attacker non-deterministically opens as many new
sessions as it wants and the causality structure in which this scenario
(protocol + attacker) has been defined, can be analyzed for a state that is
considered erroneous with respect to some security requirements decided a
priori. Therefore, as we mentioned in Section~\ref{sec:problem}, the
formal verification of the security of protocols can be considered a scientific
security theory; the implementations resulting from the formalized (and
verified) design can be tested against the formalized security requirements.
So, it is not completely true that ``Thus, formal approaches offer no escape
from our basic problem; only by making unfalsifiable assumptions (about what an
attacker can do) will they allow derivation of a necessary condition'' (as
Herley describes in\autocite{Herley2009so}) rather, the
unfalsifiability-related issue seems to stem from how security conditions are
defined, i.e. how security requirements are defined. One can easily see that
this is not just an issue in the formal protocol verification community but the
very foundations of the concept of Weakness and Vulnerability are defined in a
way that makes it impossible (they are necessary conditions) to make a general
security statement on any system (even if we tested every single Weakness on
the CWE on our system it may be the case that a Weakness han yet to be
discovered). Therefore, we hypothesize that a security theory should define in
a complete way how errors (and then Weaknesses, Vulnerability, and Exploits)
may be introduced.

>>>>

\paragraph{Attacker as a Dis-honest Agent}
As we mentioned in Section~\ref{sec:problem} when we postulated the
security hypothesis, Weaknesses and Vulnerability are the two core security
concepts of the current understanding of what security is. However, we see two
(opposed in terms of abstraction) possibility to better defined those concepts:
defining how an agent (Mankind or Nature) can maliciously use Weaknesses to
produce Vulnerability (and, in turn, Exploits), or how Errors can be defined
(since Error is a term from which Weakness are, in turn, defined).  The former,
in our current formalization, implies detailing and constraining the Causality
Principle by introducing constraints on how agents acts. For example,
introducing the concept of intentionality, motivation, or desire (see the BDI
framework defined in \autocite{Rao1995semantics} for a general overview of those concepts).
The symbolic approach (also called Dolev-Yao model) in the formal verification
of security and communication protocols, can be used in this spirit. The
attacker can be seen as an additional agent (often identified with the network
where the messages are exchanged) that sees the network traffic and manipulates
it. The actions are generic (the attacker can combine two messages, split a
message into its components, or decrypt/encrypt if it has the proper security
key) but not something that an honest agent cannot do. The main difference is
that an honest agent follows the nominal specification of the protocol while
the attacker doesn't. The attacker non-deterministically opens as many new
sessions as it wants and the causality structure in which this scenario
(protocol + attacker) has been defined, can be analyzed for a state that is
considered erroneous with respect to some security requirements decided a
priori. Therefore, as we mentioned in Section~\ref{sec:problem}, the
formal verification of the security of protocols can be considered a scientific
security theory; the implementations resulting from the formalized (and
verified) design can be tested against the formalized security requirements.
So, it is not completely true that ``Thus, formal approaches offer no escape
from our basic problem; only by making unfalsifiable assumptions (about what an
attacker can do) will they allow derivation of a necessary condition'' (as
Herley describes in\autocite{Herley2009so}) rather, the
unfalsifiability-related issue seems to stem from how security conditions are
defined, i.e. how security requirements are defined. One can easily see that
this is not just an issue in the formal protocol verification community but the
very foundations of the concept of Weakness and Vulnerability are defined in a
way that makes it impossible (they are necessary conditions) to make a general
security statement on any system (even if we tested every single Weakness on
the CWE on our system it may be the case that a Weakness han yet to be
discovered). Therefore, we hypothesize that a security theory should define in
a complete way how errors (and then Weaknesses, Vulnerability, and Exploits)
may be introduced.

### 2020-06-05 reports/white_paper/theory.tex
minor changes

\Paragraph{Methodology}
The concept of Knowledge \autocite{Hintikka1962knowledge} (from an
epistemological point of view) can be defined as the way to get to the Truth.
Anything that is known is considered true\footnote{It is also
believed that Knowledge, defined as we did, is in-apprehensible \autocite{Empiricus1990Pyrrhonism}, but this shouldn't be an issue since it is a common problem to every scientific search}
In contrast, Beliefs \autocite{Hintikka1962knowledge} are potentially true and
leads to new hypothesis. Due to the
structure of deductive reasoning we can only start from unproven hypothesis and test the
deductive conclusions on the reality. Therefore, whatever a scientific-security-theory may be,
it can only be defined starting from hypotheses.  A sound reasoning should lead to
conclusion that shall be tested on a real system. Therefore, only at the end of
the reasoning we will be able to test, and judge, the quality of the theory
itself. The identification of hypotheses should start from probably true
beliefs (e.g. from Information \autocite{Hintikka1993Information}), and then from the state of the art.

@@

\Paragraph{Number of different configurations of a system}
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different sets, where $r^e$ is the number of permutation of $r$
relations over $e$ elements with repetitions, with $e$ being the number of
$k$-ary combinations of $n$ sets, $\binom{n}{k}$.
In our case, $\binom{n}{k}=3$ since we consider $3$ sets
($\assertionRegion,\behaviorRegion,\factRegion$), and all the relations
considered in the RCC are binary.  Hence, using RCC5 (with five different
spatial relations) over three sets, we can theoretically define up to 125
different type of agents. However, only 54 of the 125 (as showed in
\cite{improvingRCC}) combinations are topologically correct with respect to
the definition of the relations of RCC5. Generalizing to all the RCCs:

@@

\Paragraph{\Rcc{$\assertionRegion$}{$\behaviorRegion$} -- Collaboration}
In order to reason on the relation between assertions and behavior we first
need to consider that, by definition, assertions are defined as transfer of
information between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$, the Input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
two pairs of regions\footnote{``I am not asserting, as Lotze did, that a relation between $X$ and $Y$
consists of a quality in $X$ and a quality in $Y$ -- a view which I regard as
quite indefensible. -- I assert that a relation $Z$ between $X$ and $Y$ involves
the existence in $X$ of the quality ``having the relation $Z$ to $Y$'' so that
a difference of relations always involves a difference in quality, and a
change of relations always involves a change of quality.'' --  Ellis J. McTaggart, The Unreality of Time\autocite{Mctaggart1908unreality}}: 
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\behaviorRegion$},
		where the relation between Input-assertions and behavior
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation. With more details, the
		ideal specification of the functional architecture, along with
		the expected inputs, defines the functional behavior of an
		ideal system.  If all the inputs (Assertions) are correctly
		handled in the functional specification (behavior) the
		specification is complete.

@@

\Paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and \Rcc{$\behaviorRegion$}{$\factRegion$} -- Honesty and Competence}
While Assertions and Beliefs generated by the Behavior determines how a system should work, and the
relation between the two defines the quality (i.e. the completeness w.r.t the specification) of the agent, the relation of
Assertions and Behavior with Facts determines the quality with respect to the
nominal (specified) system.  Given that Facts defines what must be true in the
system, the relation of Assertions an Facts
determines the degree of quality between the information circulating in a
system (or within an agent) and reality\fixnote{mr}{assertions are reality
while facts are dogmatic assertions at specification level, i.e. requirements}.
Since the transfer of information through Assertions
generates Beliefs, a dishonest agent may circulate false information,
generating false Beliefs.  We note that it is often implied that the intention behind
circulating false information discerns a dishonest and an incompetent agent,
however, we consider honesty related to sharing truths\footnote{``[\ldots] truth
exists only in the good man, but the true in the bad man as well; for it is
possible for the bad man to utter something true.''
Sextus Empiricus, Outlines of Pyrrhonism, II-83\autocite{Empiricus1990Pyrrhonism}}.  The
relation between Beliefs and Facts determines the competence (on the subjects
defined by the Facts) of an agent (i.e. the more competent an agent is, the
more likely a Belief of that agent is true).

>>>>

\paragraph{Methodology}
The concept of Knowledge \autocite{Hintikka1962knowledge} (from an
epistemological point of view) can be defined as the way to get to the Truth.
Anything that is known is considered true\footnote{It is also
believed that Knowledge, defined as we did, is in-apprehensible \autocite{Empiricus1990Pyrrhonism}, but this shouldn't be an issue since it is a common problem to every scientific search}
In contrast, Beliefs \autocite{Hintikka1962knowledge} are potentially true and
leads to new hypothesis. Due to the
structure of deductive reasoning we can only start from unproven hypothesis and test the
deductive conclusions on the reality. Therefore, whatever a scientific-security-theory may be,
it can only be defined starting from hypotheses.  A sound reasoning should lead to
conclusion that shall be tested on a real system. Therefore, only at the end of
the reasoning we will be able to test, and judge, the quality of the theory
itself. The identification of hypotheses should start from probably true
beliefs (e.g. from Information \autocite{Hintikka1993Information}), and then from the state of the art.

@@

\paragraph{Number of different configurations of a system}
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different sets, where $r^e$ is the number of permutation of $r$
relations over $e$ elements with repetitions, with $e$ being the number of
$k$-ary combinations of $n$ sets, $\binom{n}{k}$.
In our case, $\binom{n}{k}=3$ since we consider $3$ sets
($\assertionRegion,\behaviorRegion,\factRegion$), and all the relations
considered in the RCC are binary.  Hence, using RCC5 (with five different
spatial relations) over three sets, we can theoretically define up to 125
different type of agents. However, only 54 of the 125 (as showed in
\cite{improvingRCC}) combinations are topologically correct with respect to
the definition of the relations of RCC5. Generalizing to all the RCCs:

@@

\paragraph{\Rcc{$\assertionRegion$}{$\behaviorRegion$} -- Collaboration}
In order to reason on the relation between assertions and behavior we first
need to consider that, by definition, assertions are defined as transfer of
information between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$, the Input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
two pairs of regions\footnote{``I am not asserting, as Lotze did, that a relation between $X$ and $Y$
consists of a quality in $X$ and a quality in $Y$ -- a view which I regard as
quite indefensible. -- I assert that a relation $Z$ between $X$ and $Y$ involves
the existence in $X$ of the quality ``having the relation $Z$ to $Y$'' so that
a difference of relations always involves a difference in quality, and a
change of relations always involves a change of quality.'' --  Ellis J. McTaggart, The Unreality of Time\autocite{Mctaggart1908unreality}}: 
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\behaviorRegion$},
		where the relation between Input-assertions and behavior
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation. With more details, the
		ideal specification of the functional architecture, along with
		the expected inputs, defines the functional behavior of an
		ideal system.  If all the inputs (Assertions) are correctly
		handled in the functional specification (behavior) the
		specification is complete.

@@

\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and \Rcc{$\behaviorRegion$}{$\factRegion$} -- Honesty and Competence}
While Assertions and Beliefs generated by the Behavior determines how a system should work, and the
relation between the two defines the quality (i.e. the completeness w.r.t the specification) of the agent, the relation of
Assertions and Behavior with Facts determines the quality with respect to the
nominal (specified) system.  Given that Facts defines what must be true in the
system, the relation of Assertions an Facts
determines the degree of quality between the information circulating in a
system (or within an agent) and reality\fixnote{mr}{assertions are reality
while facts are dogmatic assertions at specification level, i.e. requirements}.
Since the transfer of information through Assertions
generates Beliefs, a dishonest agent may circulate false information,
generating false Beliefs.  We note that it is often implied that the intention behind
circulating false information discerns a dishonest and an incompetent agent,
however, we consider honesty related to sharing truths\footnote{``[\ldots] truth
exists only in the good man, but the true in the bad man as well; for it is
possible for the bad man to utter something true.''
Sextus Empiricus, Outlines of Pyrrhonism, II-83\autocite{Empiricus1990Pyrrhonism}}.  The
relation between Beliefs and Facts determines the competence (on the subjects
defined by the Facts) of an agent (i.e. the more competent an agent is, the
more likely a Belief of that agent is true).
